The evolution of gene expression in the term placenta of viviparous mammals

### 2015-10-08 manuscript/manuscript.Rnw
move acronyms to their own file
calculate core placenta transcriptome

% Acronyms
\acrodef{EV}[EV]{Extracellular Vesicle}
\acrodef{SVM}{Support Vector Machine}
\acrodef{ROC}{Receiver Operating Characteristic}
\acrodef{KNN}{K Nearest Neighbors}
\acrodef{SRA}{Sequence Read Archive}
\acrodef{FPKM}{Fragments Per Kilobase of exon per Million reads}

% \twocolumn
\maketitle
\onehalfspacing

@@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees"))) {
    load(a)
}
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
@

@@

>>>>

% Acronyms
\input{acronyms}
% \twocolumn
\maketitle
\onehalfspacing

@@

<<load.data,echo=FALSE,results="hide",cache.extra=tools::md5sum(paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset")))>>=
for (a in paste0("../data/",c("combined_fpkm","gene_trees","housekeeping_genes_superset"))) {
    load(a)
}
setkey(genes.to.tree,"gene_id")
star.logs[,value:=gsub("%$","",value)]
star.logs.wide <- 
    data.table(dcast(star.logs,species+file~field))
setnames(star.logs.wide,
         "Average input read length",
         "input.read.length")
setnames(star.logs.wide,
         "Number of input reads",
         "input.reads")
setnames(star.logs.wide,
         "Uniquely mapped reads number",
         "unique.map.reads")
@

@@

\subsection{Core Placenta Transcriptome}

\begin{figure}
<<core.placenta.transcriptome,fig.width=8,fig.height=12,echo=FALSE>>=
combined.fpkm.wide <- 
    data.table(dcast(combined.fpkm,
                     human_name~species,
                     fun.aggregate=sum,
                     value.var="mean_fpkm"))
temp <- 
    combined.fpkm.wide[,median_expression := apply(combined.fpkm.wide[,-1,with=FALSE],1,median)]
temp <- 
    combined.fpkm.wide[,min_expression := apply(combined.fpkm.wide[,-1,with=FALSE],1,min)]
setkey(combined.fpkm.wide,"human_name")
core.placenta.transcriptome.long <- 
    data.table(melt(combined.fpkm.wide[min_expression >= 10,
                                       ][order(-median_expression)
                                         ][!(human_name %in% 
                                                 housekeeping.genes.superset[,gene_short_name])],
                    id.vars="human_name",
                    variable.name="Species",
                    value.name="FPKM"
                    ))
temp <- core.placenta.transcriptome.long[,human_name:=
                                     reorder(human_name,
                                             combined.fpkm.wide[core.placenta.transcriptome.long[,human_name],median_expression])]
rm(temp)

print(ggplot(core.placenta.transcriptome.long[!Species %in% c("median_expression","min_expression"),],
             aes(y=human_name,x=capfirst(Species)))
      + geom_tile(aes(fill=log10(FPKM)))
      + scale_x_discrete(expand=c(0,0))
      + scale_y_discrete(expand=c(0,0))
      + scale_fill_gradient(low="white",
                            high="blue",
                            limits=c(0,4),
                            guide=guide_colorbar(title=expression(log[10](FPKM)))
                            )
      + ylab("Human Ortholog Name")
      + xlab("Species")
      + theme_grey()
      + theme(legend.position="top",
              axis.ticks = element_blank(), 
              axis.text.x = element_text(angle = 340, hjust = 0.3, vjust=0.8,face="italic"))
      )

# 
# print(xtable(combined.fpkm.wide[min_expression > 10,
#                                 ][order(-median_expression)
#                                   ][!(human_name %in% 
#                                           housekeeping.genes.superset[,gene_short_name])][1:75,],
#              digits=0
#              ),
#       include.rownames=FALSE,
#       booktabs=TRUE,
#       floating=FALSE
#       )
@ 
\caption{Core Placenta Transcriptome.
  \Sexpr{core.placenta.transcriptome.long[,length(unique(human_name))]}
  genes expressed (\ac{FPKM}$\ge 10$) in the placenta of all organisms
  studied with housekeeping
  genes\cite{Eisenberg.ea2013:Humanhousekeepinggenesrevisited}
  removed.}
\label{fig:core_placenta_transcriptome}
\end{figure}

Annexins?

TIMP3 -- maybe in pre-eclampsia (23172037)
SERPINH1? with TIMP3 
% http://www.proteinatlas.org/ENSG00000149257-SERPINH1/tissue/placenta

Why are RPL17 and RPS26 here? They are ribosomal proteins. These
should be missing now.
